---
title: "rpostgisLT test 2017"
author: "Balázs Dukai"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_notebook
---

```{r setup, message=FALSE}
library(rpostgis)
library(lubridate)
library(rpostgisLT)
library(sf)
library(shiny)
library(miniUI)
library(leaflet)
```

# Part 1

Read roe deer trajectories

```{r read_roe_gps}
data("roe_sensors_animals_tables")
data("roe_gps_data")

gps_data <- rbind(roe_gps_data$GSM01511[c("gps_sensors_code","utc_date",
                                          "utc_time","latitude","longitude")],
                  roe_gps_data$GSM01512[c("gps_sensors_code","utc_date",
                                        "utc_time","latitude","longitude")])
```

Create an `ltraj` from the data. It is important to set the correct time zone and the Coordinate Reference System (CRS) of the relocations. The time zone is *UTC* as it is denoted in the variable names. There is no explicit information on the CRS, but because the relocations are raw GPS measurements, they are most likely in *EPSG:4226* which is the CRS used by GPS.

```{r ltraj_roe}
gps_data$tstamp <- dmy_hms(paste(gps_data[, "utc_date"],
                                 gps_data[, "utc_time"], sep = " "),
                           tz = "UTC")
gps_data <- dplyr::rename(gps_data, x = longitude, y = latitude)
gps_data <- dplyr::select(gps_data, x, y, tstamp, gps_sensors_code)

proj4string <- sp::CRS("+init=epsg:4326")

roe_ltraj <- adehabitatLT::as.ltraj(gps_data, date = gps_data$tstamp,
                                    id = gps_data$gps_sensors_code,
                                    proj4string = proj4string)
```

Thest the `hab::trajdyn()` function.

```{r trajdyn, message=FALSE}
hab::trajdyn(roe_ltraj)
```

Put the *ltraj* into a database.

```{r to_db}
ltraj2pgtraj(conn, roe_ltraj, schema = "test2017_1")
```

# Part 2

For the Shiny App the trajectories are directly pulled from the database and the app uses Leaflet to plot the trajectories on a base map. By using a reactive expression it is possible to interactively subset the *steps* based on their time stamp.

```{r shiny}
subsetSteps <- function(conn) {
    # get trajectories from the database
    SQL <- paste0("select step_id, step_geom, relocation_time
                  from test2017_1.step_geometry_roe_ltraj
                  where step_geom is not null")
    x <- suppressWarnings(st_read_db(conn, query = SQL))
    minT <- min(x$relocation_time)
    maxT <- max(x$relocation_time)
    
    ui <- bootstrapPage(
        
      tags$style(type = "text/css", "html, body {width:100%;height:100%}"),
      
      leafletOutput("map", width = "100%", height = "100%"),
      
      absolutePanel(top = 10, right = 10,
                    sliderInput("range", "Timestamp", minT, maxT, 
                                value = c(minT, maxT), step = 60
                                )
                    )
    )
    
    server <- function(input, output, session) {
        
        # reactivbe expression
        filteredData <- reactive({
          x[x$relocation_time >= input$range[1] & x$relocation_time <= input$range[2], ]
        })
        
        # Leaflet base map, and starting view centered at the trajectories
        output$map <- renderLeaflet({
            leaflet(x) %>%
                addTiles() %>% 
                fitBounds(lng1 = 10.9872122, lat1 = 45.9882219,
                          lng2 = 11.0587093, lat2 = 46.0467987)
        })
        
        # update plotted trajectories with the reactive expression output
        observe({
            leafletProxy("map", data = filteredData()) %>%
                clearShapes() %>%
                addPolylines(
                    fillOpacity = 1,
                    opacity = 1,
                    color = "#de2d26",
                    weight = 3
                )
        })
        
    }
    
    shinyApp(ui, server)
}
```

```{r}
subsetSteps(conn)
```

# Part 3

Read required data.

```{r read_for_db}
data("roe_gps_data")
gps_data <- roe_gps_data$GSM01511[c("utc_date","utc_time","latitude","longitude")]

data("roe_raster")
raster <- roe_raster$srtm_dem
```

Set up database for storing trajectories.

```{r db_setup}
pgtrajSchema(conn, schema = "test2017")
RPostgreSQL::dbWriteTable(conn, name = "gps_data", value = gps_data)
rpostgis::pgWriteRast(conn, name = c("test2017", "roe_raster"), raster = raster)
```

Parse time stamps, including correct time zone (UTC), and prepare bursts from year-month combinations. Update the raw data table with the parsed values.

The RS DBI driver complains that the result set does not correspond to a SELECT statement, which is expected. Nevertheless the process runs through.

```{sql connection=conn}
ALTER TABLE
    public.gps_data ADD tstamp TIMESTAMPTZ;

ALTER TABLE
    public.gps_data ADD burst TEXT;
    
ALTER TABLE
    public.gps_data ADD rid INTEGER;

UPDATE
    public.gps_data
SET
    tstamp = a.t,
    rid = a.id::integer
FROM
    (
        SELECT
            "row.names" AS id,
            (
                split_part(
                    utc_date,
                    '/',
                    3
                )|| '-' || split_part(
                    utc_date,
                    '/',
                    2
                )|| '-' || split_part(
                    utc_date,
                    '/',
                    1
                )|| ' ' || utc_time || ' ' || 'UTC'
            )::TIMESTAMPTZ AS t
        FROM
            public.gps_data
    ) a
WHERE
    public.gps_data."row.names" = a.id;

UPDATE
    public.gps_data
SET
    burst = a.b
FROM
    (
        SELECT
            "row.names" AS id,
            SUBSTRING( tstamp::TEXT, 1, 7 ) AS b
        FROM
            public.gps_data
    ) a
WHERE
    public.gps_data."row.names" = a.id;
```

Create *pgtrajes* from the raw data.

```{r}
as_pgtraj(conn, c("public", "gps_data"), schema = "test2017",
          relocations = c("longitude", "latitude"), timestamps = "tstamp",
          bursts = "burst", srid = 4326, rids = "rid")
```

By using spatial intersection on the trajectories and the elevation raster, the average elevation values for each step are extracted.

```{sql connection=conn}
SELECT
    step_id,
    AVG( ST_Value( rast,( gd ).geom )) z
FROM
    (
        SELECT
            step_id,
            ST_DumpPoints(step_geom) AS gd
        FROM
            test2017.step_geometry_pgtraj
        WHERE
            step_geom IS NOT NULL
    ) AS traj
LEFT JOIN test2017.roe_raster ON
    ST_Intersects(
        rast,
        (gd).geom
    )
GROUP BY
    step_id;
```

